---
title: 'MAT 217: Homework 1'
author: "Your name"
date: "`r Sys.Date()`"
output: 
  openintro::lab_report: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(hanoverbase)
```

# INTRODUCTION 

The main objective of this homework is to help you get used to RStudio interface and its functionalities. We will use the dataset `in09_house_district_summary.csv` that you might find in the `data` folder in the directory throughout this homework. The dataset is renamed into `district` for this homework. 

```{r import dataset, message=FALSE, warning=FALSE}
district <- read_csv("../../data/in09_house_district_summary.csv")
```

# Question 1 

## Part 1a

How many cases (or rows) are there in this dataset? 

> ANSWER

## Part 1b

What does one case represent? 

> ANSWER

# Question 2

## Part 2a

How many variables are there in the dataset? You may either count manually or use the `dim()` function in the code block below.   

> ANSWER

```{r}

```

## Part 2b

Which variables are quantitative? Indicate if they are discrete or continuous. 

> ANSWER

## Part 2c

Which variables are categorical? For each variable, list out the corresponding levels (categories). 

> ANSWER: 

## Part 2d

Is the variable `year` best treated as a quantitative or a categorical variable? Explain your reasoning. 

> ANSWER: 

# Question 3

Run the following code to plot the graph

```{r}
plot(district$year, district$total_votes, 
     main = "Total votes 2018 - 2024", 
     xlab = "Year", 
     ylab = "Total votes")
```

State your observations from the graph: 

1. Identify the year with the lowest total votes. 
2. Identify the year with the highest total votes. 
3. Briefly describe how the total votes change over the years.

Do not attempt to explain why such pattern occurred. We will go into the reasoning in the subsequent homework. 

> ANSWER: 

# OPTIONAL EXPLORATION

The vote margin, stored in the variable `rep_margin`, is defined as:

<center> 

rep_margin = rep_votes - dem_votes

</center>

A positive difference means that the Republican candidate received more votes. A negative difference means that the Democratic candidate received more votes. 

Modify the code in Question 3 to display the plot of voting margin over the years. Your codes should be in the following code block.

```{r}

```

What pattern do you notice from the plot? 

# Question 4

1. Knit the document and submit the HTML file on Canvas. 

2. Remember to record your individual reflection from this homework in the observation-log. 

